### 2023-05-03 paper.Rmd
minor

To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering. K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean. Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis. After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming the traditional anticlustering criteria diversity and k-means. Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].

@@

Anticlustering is usually implemented by reversing an objective function used in classical cluster analysis. Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups. While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data, e.g., by defining a taxonomy on naturally occurring groups [@everitt1979]. Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern. Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similar.  A good clustering solution assigns similar elements into the same cluster and dissimilar elements into different clusters. A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This distinction can be clarified using Figure 1 where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions. An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution. Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not. In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%. Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions. Interestingly, the k-plus objective introduces an additional distinction between cluster analysis and anticlustering. Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering. Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.

>>>>

To optimize overall between-group similarity---and hence to overcome the limitations of k-means anticlustering---this paper introduces k-plus anticlustering. K-plus is an extension of the k-means objective that quantifies between-group similarity as discrepancy with regard to several distribution characteristics instead of only the mean. Specifically, k-plus offers the possibility to minimize differences with regard to the variance and higher order moments such as skewness and kurtosis. After formally introducing the k-plus objective, two simulation studies show that k-plus anticlustering is well-suited to create groups having minimal differences according to multiple distribution characteristics, significantly outperforming traditional anticlustering criteria such as the diversity and the classical k-means objective. Practical examples illustrate how readers can easily apply k-plus anticlustering using the free and open source software package `anticlust` [@papenberg2020; @R-anticlust], an extension to the popular statistical programming language R [@R-base].

@@

Anticlustering is usually implemented by reversing an objective function used in classical cluster analysis. Whereas cluster analysis seeks homogeneity within groups and separation between groups, the reversal results in heterogeneity within groups and similarity between groups. While tightly linked in a mathematical sense, an important semantic difference is that cluster analysis is concerned with finding "true" structure in the data, e.g., by defining a taxonomy on naturally occurring groups [@everitt1979]. Anticlustering instead imposes structure on the data, while the anticlusters themselves do not represent any underlying "true" pattern. Correspondingly, different---but similarly adequate---anticlustering partitions tend to look very different, while good clustering solutions tend to look similar. A good clustering solution assigns similar elements into the same cluster and dissimilar elements into different clusters. A good anticlustering solution does not depend on assigning individual items either to the same or a different group, independent of their pairwise dissimilarity.[^dispersion] This pattern was observed in SImulation 2, where it was shown that partitions obtaining high between-group similarity according to different criteria were not at all similar; the ARI was usually close to 0. For different clustering methods, there is usually a considerable amount of overlap between partitions. This  distinction between cluster analysis and anticlustering can also be clarified using Figure 1, where we observe a high concentration of suitable anticlustering partitions but only few suitable clustering solutions. An astonishing `r mean(df$kmeans / max(df$kmeans) >.999) * 100`% of all partitions are within 0.1% of the global maximum in $\mathit{SSE}$, i.e., the best k-means anticlustering solution. Among these `r sum(df$kmeans / max(df$kmeans) >.999)` partitions, every possible pair of elements is sometimes assigned to the same group and sometimes not. In contrast, there are only `r sum(min(df$kmeans)/ df$kmeans > .90)` partitions (`r mean(min(df$kmeans)/ df$kmeans > .90)*100`%) that approximate the global minimum in $\mathit{SSE}$ (i.e., the best k-means clustering solution) by even 10%. Thus, even though the problems are based on the same objective function, cluster analysis and anticlustering exhibit a stark discrepancy with regard to the solution space of partitions. Interestingly, the k-plus objective introduces an additional distinction between cluster analysis and anticlustering. Reversing the k-plus anticlustering objective no longer yields a useful clustering interpretation, because identifying groups that are different with regard to higher order moments does not constitute a clustering. Thus, even though k-plus extends the classical k-means clustering criterion, it is a pure anticlustering objective.
